\section{Introduction}
\label{sec:introduction}

\subsection{Philosophical and Physical Motivation}
The question ``Why is there something rather than nothing?'' has persisted across philosophy and physics for millennia. Classical physics permits a stable vacuum state of absolute nothingness, yet quantum mechanics reveals pervasive fluctuations, suggesting a non-zero baseline. The Origin Axiom addresses this tension by asserting that perfect global cancellation is structurally forbidden in the universe's configuration space.

Formally, consider a complex scalar amplitude $A(\mathcal{C})$ defined over the space of all possible configurations $\mathcal{C}$. The axiom enforces a strict lower bound:
\begin{equation}
|A(\mathcal{C})| > \epsilon > 0,
\end{equation}
where $\epsilon$ is a minimal positive floor (simulated as $\sim 10^{-12}$ in normalized units). This floor prevents $|A| \to 0$, rendering absolute nothingness impossible.

Philosophically, the axiom implies existence is not contingent but inherent — a structural bias against nothingness that favors complexity and differentiation. Intellectually, it offers a principled alternative to fine-tuning arguments: rather than invoking anthropic selection or multiverses, the axiom posits that the universe must exhibit non-zero structure at all scales.

\subsection{The Unifying Phase $\thetastar$}
To ground the axiom in empirical physics, we introduce a unifying phase $\thetastar$ extracted from Standard Model flavor observables. $\thetastar$ is derived from the PMNS neutrino mixing matrix and CKM quark mixing matrix, yielding a fiducial value of 3.63 radians with a robust band of [2.18, 5.54] radians (determined from joint fits in the companion repository \texttt{origin-axiom-theta-star}).

This phase $\thetastar$ serves as the bridge between flavor phenomenology and scalar dynamics, modulating amplitudes across scales. Simulations demonstrate that $\thetastar$-dependent effects propagate consistently from neutrino masses to vacuum energy, cosmological expansion, and microstructure.

\subsection{Unification Roadmap}
We demonstrate the axiom's unifying power through a systematic chain of simulations:

\begin{itemize}
    \item \textbf{Flavor bridge} (Section \ref{sec:flavor}): $\thetastar$-modulated seesaw mechanism reproduces PDG neutrino masses.
    \item \textbf{Vacuum shift} (Section \ref{sec:vacuum}): Microcavity models show $\sim$2.2\% modulation in energy shift $\Delta E(\thetastar)$.
    \item \textbf{Cosmological expansion} (Section \ref{sec:cosmology}): FRW evolution with $\Lambda(\thetastar)$ yields $\sim$1\% acceleration in scale factor.
    \item \textbf{Microstructure} (Section \ref{sec:microstructure}): 3D lattices with defects exhibit $\thetastar$-dependent stability.
    \item \textbf{Synthesis} (Section \ref{sec:synthesis}): Full chain reveals aligned patterns and emergent phenomena.
\end{itemize}

The framework is fully reproducible via open-source code at \url{https://github.com/originaxiom/origin-axiom} and \url{https://github.com/originaxiom/origin-axiom-theta-star}, with git-tagged versions ensuring traceability.

\subsection{Contributions and Scope}
This work is an exploratory yet rigorous proposal, developed collaboratively with AI assistance. As a non-traditional contributor, the author prioritizes clarity, reproducibility, and intellectual honesty. All results stem from systematic simulations; no claims are made beyond what is demonstrated. The paper aims to invite scrutiny, extension, and collaboration from the scientific community.
