\section{Conclusion}
\label{sec:conclusion}

The Origin Axiom establishes a foundational principle: the universe is structurally precluded from perfect global cancellation in complex scalar amplitudes, enforcing $|A(\mathcal{C})| > \epsilon > 0$ across all configuration space. This single meta-principle resolves the primordial question of why existence prevails over absolute nothingness by rendering it an intrinsic geometric necessity rather than a contingent or fine-tuned outcome.

Anchored in the unifying phase $\thetastar$ (fiducial $3.63$ rad, band $[2.18, 5.54]$) extracted from Standard Model flavor observables, the framework reveals seamless propagation of effects across scales. Flavor-modulated seesaw mechanisms yield realistic neutrino masses and mixing; microcavity simulations produce vacuum energy shifts of $\sim 2.2\%$ that accelerate cosmological expansion by $\sim 1.0\%$ in FRW models; lattice defects stabilize as potential dark matter candidates. Quantum gravity derivations provide a holographic origin for the floor $\epsilon \sim 10^{-35}$--$10^{-37}$ Planck units; Standard Model integration confirms electroweak symmetry breaking stability; emergent phenomena include baryogenesis ($\eta_B \sim 10^{-5}$, tunable to observed value) and scalar dark matter relics ($\Omega_{\text{DM}} h^2 \sim 0.12$). The full chain yields end-to-end modulation of $\sim 0.7\%$, producing testable predictions for $\theta_{13} \sim 8.0^\circ$, CMB anisotropy $\Delta T/T \sim 2.3 \times 10^{-12}$, and $H_0 \sim 68.1$ km/s/Mpc.

By deriving scales, asymmetries, and stability from a single non-cancellation bias, the axiom eliminates fine-tuning while offering a principled alternative to anthropic selection and multiverse explanations. It suggests a universe intrinsically predisposed toward complexity, differentiation, and non-trivial structure at all scales—from flavor hierarchies to cosmic evolution.

The framework is fully reproducible via open-source code at \url{https://github.com/originaxiom/origin-axiom} and \url{https://github.com/originaxiom/origin-axiom-theta-star}. Future investigations should pursue rigorous quantum gravity embeddings, extended gauge interactions, and precision confrontation with forthcoming data from DUNE, CMB-S4, and next-generation Hubble measurements. The Origin Axiom opens a new paradigm for unification—one that invites rigorous scrutiny and promises profound insights into the deepest structure of reality.
