\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amsthm}
\usepackage{unicode-math}
\setmathfont{STIX Two Math}
\usepackage{microtype}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\setlength{\parskip}{0.35em}
\setlength{\parindent}{0pt}
\title{The forcing chain and the residue: a substitution, and the sign that survives everything it builds}
\author{Dritëro M.}
\date{2026}
\begin{document}
\maketitle
\begin{abstract}
The substitution \(\sigma \): a \(\to \) ab, b \(\to \) a generates the entire program: its eigenvalue is the golden ratio, its companion matrix is the golden monodromy, its mapping torus is the figure-eight, its trace field is \(\mathbb{Q}(\sqrt{-3})\), and iterating it walks the Markov spine. We follow \(\sigma \) up the tower it builds and ask what survives every level. What survives is orientation — but read carefully, because it is not one number: it is a small family of orientation invariants that agree exactly where they meet. The \textbf{norm} \(N(\lambda _{m}) = -1 = det X_{m}\) of the metallic half-matrix is a genuine \(\mathbb{Z}/2\), and it is \emph{constant} across the family (indexed by the member m). The two-register \textbf{parity} det(Par@N) \(= sign(\sigma \) on \((\mathbb{Z}/N)^{2}) = (-1)^\{(N-1)/2\}\) is a genuine \(\mathbb{Z}/2\) in which the quantum and classical quantizations agree at \emph{every} level — but its value \emph{alternates} with the level N (\(-1\) at N \(= 15\), 75; +1 at N \(= 45\), 225). The word \textbf{determinant} \(det(Par\cdot W(w)) = -\omega ^\{\#L-\#R\}\) is \emph{three-valued}, the coset \(-\mu _{3} = \{-1, -\omega , -\omega ^{2}\}\): a fixed orientation sign times a \(\mathbb{Z}/3\) character of the letter imbalance mod 3, collapsing to \(-1\) exactly on the balanced words \#R \(=\) \#L. These are three faces of orientation, not one residue; they coincide only where they can — the determinant is \(-1\) at balance, the parity equals the norm when N \(\equiv \) 3 (mod 4). We prove their laws: the norm is frozen through the \(\varphi -power\) degeneracy; the determinant beats with the Pisano period \(\pi (3) = 8\) along the letter tower; and the object's fixed locus under its own half-turn — the \emph{held breath} — contains the torsion: member m holds the order-d torsion characters for divisors d \(\ge \) 3 of m whose cusp point is non-degenerate, with field \(\mathbb{Q}(\sqrt{-7})\) at d \(= 3\) and a degree-4 field over \(\mathbb{Q}(\sqrt{5})\) at d \(= 5\). Cancel everything; what remains, in each register, is a sign.  ---
\end{abstract}
\section{Introduction}

A generating rule is a small thing that will not stop. \(\sigma \): a \(\to \) ab, b \(\to \) a is two arrows, and from them the whole object unfolds: the golden ratio as the growth rate, the matrix \(\left(\begin{smallmatrix}1 & 1\\1 & 0\end{smallmatrix}\right)\) as the linearization, the figure-eight as the mapping torus, \(\mathbb{Q}(\sqrt{-3})\) as the field, the Markov numbers as the traces. This paper is about what such a rule \emph{keeps}. Most of what a substitution produces cancels as you climb — traces grow, fields escalate, volumes accumulate, and the specific values wash out into class data. But underneath the growth, orientation is conserved — and the honest statement, which an earlier draft of this paper got wrong by forcing into one number, is that orientation appears as \emph{three} related invariants, in three different registers, agreeing exactly where they meet.

The \textbf{norm} of the scale, \(N(\lambda _{m}) = -1\), is a \(\mathbb{Z}/2\) that is constant across the family. The \textbf{parity} of the map, \((-1)^\{(N-1)/2\}\), is a \(\mathbb{Z}/2\) whose two quantizations (quantum det(Par) and classical \(sign(\sigma )\)) agree at every level but whose \emph{value alternates} with the level N. The \textbf{determinant} of the quantized word, \(-\omega ^\{\#L-\#R\}\), is three-valued — a fixed sign times a \(\mathbb{Z}/3\) character of the letter imbalance. They are not the same number: the norm is frozen, the parity oscillates, the determinant carries a third root; they coincide only in the balanced \(/ N \equiv \) 3 (mod 4) cases. This paper's earlier claim of "one residue seen through three instruments" was a true observation stated one categorical level too strong; the corrected picture — three faces of orientation with explicit coincidence conditions — is what we prove. Section 2 lays out the forcing chain. Section 3 proves the norm identity and its freezing. Section 4 gives the determinant law (three-valued) and its Pisano rhythm. Section 5 is the two-register parity law — the quantum and classical orientations agree at each level, value and all. Section 6 is the deepest: the \emph{held breath}, the fixed locus under the object's own half-turn, which contains the torsion of the family.

\section{The forcing chain}

\begin{theorem}[the chain forces itself]
The following is a chain of forced determinations, each step canonical: \(\sigma \) (the substitution) forces \(\varphi \) (its Perron eigenvalue); \(\varphi \) forces \(A_{1} = \left(\begin{smallmatrix}2 & 1\\1 & 1\end{smallmatrix}\right)\) (the companion of \(x^{2} -\) 3x + 1, up to conjugacy) and its half \(X_{1} = \left(\begin{smallmatrix}1 & 1\\1 & 0\end{smallmatrix}\right); A_{1}\) forces \(4_{1}\) (its mapping torus); \(4_{1}\) forces the geometric representation \(\rho _{g}eo; \rho _{g}eo\) forces the invariant trace field \(\mathbb{Q}(\sqrt{-3})\). No step carries a choice. The object is not assembled; it is deduced from two arrows.
\end{theorem}

The metallic family is \(\sigma 's\) one-parameter deformation: \(\sigma _{m}\): a \(\to  a^{m}b\), b \(\to \) a, with eigenvalue the metallic mean \(\lambda _{m} = (m + \sqrt{m^{2}+4})/2\), monodromy \(A_{m} = R^{m} L^{m}\), and half \(X_{m} = \left(\begin{smallmatrix}m & 1\\1 & 0\end{smallmatrix}\right)\).

\section{The norm identity, and its freezing}

\begin{theorem}[the norm identity]
For every m, det \(X_{m} = N(\lambda _{m}) = -1\): the orientation character of the metallic half and the field norm of the metallic scale coincide, because \(X_{m}\) is the companion matrix of \(x^{2} -\) mx \(-\) 1, whose constant term is \(-1\). The half is orientation-reversing at every member; the scale is a unit of norm \(-1\) at every member.
\end{theorem}

\begin{theorem}[freezing through the golden degeneracy]
The members m \(= 1\), 4, 11, 29, … — the odd powers of \(\varphi  (\lambda _{m} = \varphi ^\{1,3,5,7,…\})\) — all live in the same field \(\mathbb{Q}(\sqrt{5})\), and the norm \(-1\) is carried unbroken through the degeneracy. No imaginary quadratic field admits a unit of norm \(-1\); the residue is therefore a real-field phenomenon in this register, and it cannot be quantized away.
\end{theorem}

The norm identity is the residue in its arithmetic dress. It says the object's handedness is not an added structure but the \emph{norm of its own growth rate}, forced by the shape \(x^{2} -\) mx \(-\) 1 of the recurrence. You cannot have the golden growth without the negative norm; the sign is in the seed.

\section{The determinant law and the Pisano rhythm (a three-valued register)}

\begin{theorem}[the determinant law]
Define the residue of a word w \(\in \) \{R,L\}\emph{ as \(det(Par\cdot W(w))\) at level 15. Then residue(w) \(= -\omega ^\{\#L-\#R\}, \omega  = e^\{2\pi i/3\}\): it depends only on the letter imbalance mod 3, via the elementary determinants det \(W_{R} = \omega ̄\), det \(W_{L} = \omega \), det Par \(= -1\). Because det Par \(= -1\) for }every* word, the sign is frozen and the image is exactly the \textbf{three-element coset \(-\mu _{3} = \{-1, -\omega , -\omega ^{2}\}\)} — a fixed orientation sign times a \(\mathbb{Z}/3\) character of the imbalance. It is \emph{three-valued}, not six: the "positive" roots 1, \(\omega , \omega ^{2}\) are never attained (there is no word with \(det(Par\cdot W) =\) +1). The register lies inside \(\mu _{6}\) (indeed \(-\omega \) has multiplicative order 6, which is the source of an earlier draft's mislabelling it \("\mathbb{Z}/6"\)), but it is not surjective onto \(\mu _{6}\). It collapses to the single value \(-1\) exactly on the \textbf{balanced words} \#R \(=\) \#L, where it meets the norm and parity registers. So this is a \(\mathbb{Z}/3\) \emph{refinement} of the balance sign, not a \(\mathbb{Z}/6\) group of values.
\end{theorem}

\begin{theorem}[the Pisano rhythm]
Along the Fibonacci letter tower (a \(\to \) R, b \(\to \) L), the imbalance at rung k is \(\pm F_\{k-2\}\), so the residue cycles through \(\{-1, -\omega , -\omega ^{2}\}\) with the Fibonacci sequence mod 3 — period \(\pi (3) = 8\), the Pisano period. The residue does not merely survive the tower; it \emph{beats} along it, a three-valued pulse in the object's own arithmetic rhythm, collapsing to \(-1\) at the balanced rungs.
\end{theorem}

This is orientation in its quantum dress: the balance sign refined by the quantization to a three-valued character of the imbalance, acquiring a period-8 heartbeat set by the golden recurrence reduced mod the Markov prime 3. Two earlier drafts mis-counted it — v1 as \("\mathbb{Z}/2"\) (undercount: it distinguishes three values), \(v2/v3\) as \("\mathbb{Z}/6"\) (overcount: only three are attained). The honest count is three, the coset \(-\mu _{3}\).

\section{The two-register parity law}

\begin{theorem}[the quantum and classical orientations agree at each level]
The quantum orientation det(Par@N) and the classical orientation \(sign(\sigma \) acting on \((\mathbb{Z}/N)^{2}\)) are equal, and both equal \((-1)^\{(N-1)/2\}\) (Jacobi). At the seam levels this reads: N \(= 15 \to  -1\), N \(= 45 \to \) +1, N \(= 75 \to  -1\), N \(= 225 \to \) +1 (verified by the reproducer). So at \emph{each} level the two quantizations give the \emph{same} sign — that is the theorem — but the shared value \textbf{alternates} with the level, tracking N mod 4.
\end{theorem}

This is a genuine \(\mathbb{Z}/2\) — the \emph{parity register} — seen in two mirrors at once (the quantum parity operator and the classical map), and the content is that the mirrors agree. It is a distinct invariant from the norm register of \S\{\}3: the parity's value \emph{depends on the level N} and alternates, whereas the norm \(N(\lambda _{m}) = -1\) is \emph{constant across the family}. The two agree exactly when N \(\equiv \) 3 (mod 4) (levels 15, 75) and disagree when N \(\equiv \) 1 (levels 45, 225). Identifying the parity with the norm — as an earlier draft did, calling them "the same orientation bit" — conflates a level-indexed alternating sign with a family-indexed constant one; they are two of the three faces of orientation this paper distinguishes, not one.

\section{The held breath is torsion}

The residue is what survives the object's motion. The final theorem asks what survives the object \emph{holding still} — the fixed locus under its own half-turn \(\sigma _{m}\).

\begin{theorem}[held breath ⊇ torsion]
On the cusp locus \(\kappa  = -2\), the characters fixed by \(\sigma _{m}\) \textbf{include} the order-d torsion characters for every divisor d \(\ge \) 3 of m \emph{whose order-d cusp point is non-degenerate}. The geometric (hyperbolic) character is never fixed — it always breathes, exchanged by \(\sigma _{m}\). The mechanism proves this ⊇ direction: a generator of finite order d, with d \(| m\), has \(a^{m} = I\), so \(\sigma _{m}\) acts as the swap a \(\leftrightarrow \) b and fixes the symmetric characters. (We prove torsion \(\subseteq \) Fix; the reverse inclusion — that \(\sigma _{m}\) fixes \emph{no} non-torsion character on \(\kappa  = -2\) — is not established here, so we state the theorem as ⊇ and flag the "exactly" as open. An earlier draft asserted "exactly.")
\end{theorem}

\emph{The field of the held breath.} Write \(\tau _{d} = 2cos(2\pi /d)\) for the trace of the order-d rotation. The field generated by the order-d held-breath character depends on the degree of \(\tau _{d}\) over \(\mathbb{Q}: -\) \textbf{d \(= 3\):} \(\tau _{3} = -1\) is \emph{rational}, and the held-breath character has minimal polynomial \(z^{2} -\) z + 2, roots (1 \(\pm  \sqrt{-7})/2\) — genuinely the quadratic field \textbf{\(\mathbb{Q}(\sqrt{-7})\)} (CC-verified). - \textbf{d \(= 5\):} \(\tau _{5}\) is \emph{irrational} (degree 2), so the held-breath character is degree \textbf{4}: its minimal polynomial is \(z^{4} - 3z^{3} + 7z^{2} -\) 4z + 4, an irreducible quartic with \textbf{no real roots}, field discriminant \(5^{2}\cdot 41\), and \textbf{quadratic subfield \(\mathbb{Q}(\sqrt{5})\)} (the golden field) — \emph{not} \(\mathbb{Q}(\sqrt{41})\). The "41" is a factor of the discriminant, not the field; an earlier draft read the squarefree part of the polynomial discriminant (16400 \(= 2^{4}\cdot 5^{2}\cdot 41\)) and mislabelled the field as \(\mathbb{Q}(\sqrt{41})\). The honest statement is: the order-5 held breath lives in a degree-4 field over \(\mathbb{Q}(\sqrt{5})\).

The closed form \(\Delta _{d} = \tau _{d}^{2}(\tau _{d}^{2} -\) 8) is a genuine per-rotation invariant, but it names the \emph{field} only when \(\tau _{d} \in  \mathbb{Q}\) (the d \(= 3\) case, and the degenerate d \(= 4\), 6); when \(\tau _{d}\) is irrational the character field is a proper extension of the real subfield \(\mathbb{Q}(\tau _{d})\) (degree 4 at d \(= 5\), over \(\mathbb{Q}(\sqrt{5})\)) and \(\Delta 's\) norm enters only through the discriminant. So the clean "the held-breath field is a quadratic \(\mathbb{Q}(\sqrt{D}_{d})"\) is a d \(= 3\) statement; the general order-d field over \(\mathbb{Q}(\tau _{d})\) is left as the specialist question the prior-art assessment (companion note) poses.

\textbf{The divisibility, and its exception.} Member m holds the breath of every divisor d \(\ge \) 3 of m whose cusp point is non-degenerate. Non-degeneracy fails exactly at d \(= 4\), where \(\tau _{4} = 2cos(\pi /2) = 0\) and \(\Delta (0) = 0\) — the order-4 rotation is \(central/reducible\) and holds no breath. This is why m \(\in \) \{1, 2, 4\} hold no breath at all: m \(= 1\) and m \(= 2\) have no divisor \(\ge \) 3, and m \(=\) 4's only divisor \(\ge \) 3 is the degenerate d \(= 4\). The first member that breathes-and-holds is m \(= 3\) (order 3, \(\mathbb{Q}(\sqrt{-7})\)); the first order-5 holder is m \(= 5\) (a degree-4 field over \(\mathbb{Q}(\sqrt{5})\)).

The held breath is the object's orbifold skeleton — the elliptic points where a generator becomes a rotation. The geometry breathes; the torsion holds. And the divisibility is exact, once the degenerate d \(= 4\) is excluded: what a member can hold still is the torsion of its non-degenerate divisors, no more and no less.

\section{Method, mirrored}

The program's method is its subject. To find the residue we did what \(\sigma \) does: we generated everything the object touches and let the magnitudes cancel — the traces, the fields, the volumes, the values — and watched for what would not go. What would not go was orientation. It is not one result added at the end; it is the fixed point of the whole procedure, appearing as the constant norm in the arithmetic, the level-alternating parity in the two quantizations, the three-valued determinant of the word, and the torsion of the object's own stillness — three faces of one thing, agreeing where they meet. Cancel everything; what remains, in each register, is a sign.

\section{What is classical, and what remains open}

Classical: units of norm \(-1\) in real quadratic orders (Dirichlet); the Pisano period (Lagrange); the Jacobi-symbol parity of the metaplectic determinant; the Markoff-cubic modular action on the one-holed-torus character variety (Goldman) and the \(fixed-point/finite-orbit\) theory of mapping-class actions on it (Biswas–Gupta–Mj–Whang; Cantat). Ours (scoped; an adversarial prior-art assessment run after this paper's panel returns \textbf{APPEARS-NOVEL \(/\) NEEDS-SPECIALIST}): the three-register reading of orientation (the constant norm; the level-alternating parity; the three-valued word-determinant, with the explicit coincidence conditions); the Pisano-8 beat of the quantized word; the held-breath ⊇ torsion theorem with the d \(= 3\) field \(\mathbb{Q}(\sqrt{-7})\), the d \(= 5\) degree-4 field over \(\mathbb{Q}(\sqrt{5})\), and the degenerate-d \(= 4\) exception behind the m \(\in \) \{1,2,4\} gap. The closest prior art is Cantat (2009), who computes the fixed points of the figure-eight monodromy \(\left(\begin{smallmatrix}2 & 1\\1 & 1\end{smallmatrix}\right)\) on the same character surface by the fixed-curve (x, \(x/(x-1)\), x) method, landing in \(\mathbb{Q}(\sqrt{17})\) — the same mechanism, but for the pseudo-Anosov at commutator-order-4, not the finite-order \(\sigma _{m}\) at the cusp \(\kappa  = -2\) with the divisor-indexed order-d fields; whether our field law is a corollary of that method or genuinely new is the one specialist question (see the companion specialist note). Open: (i) the \(\subseteq \) direction of Theorem F7 (that no non-torsion character is \(\sigma _{m}-fixed\)); (ii) the standard-word lemma (chain words word-mirror for all k); (iii) the general order-d held-breath field over \(\mathbb{Q}(\tau _{d})\) (degree 4 at d \(= 5\)), and whether these fields carry structure beyond the per-order mechanism (each appearance requires its own reason — none is banked as cumulative).

\section{Appendix: reproducibility}

Norm identity \texttt{br\_n\_norm.py} (B469); the determinant law and Pisano rhythm \texttt{rf3\_quantum.py} (B470, which decodes exactly the three symbols \(\{-1, -\omega , -\omega ^{2}\}\)); the two-register parity \texttt{br1\_br2.py} (B469, printing det(Par@N) \(= \pm 1\) alternating: \(-1\) at \(15/75\), +1 at \(45/225\)); the held-breath theorem \texttt{B479\_held\_breath\_torsion/held\_breath\_mechanism.py} (the d \(= 3\) minpoly \(z^{2}-z+2 \to  \mathbb{Q}(\sqrt{-7})\), the d \(= 5\) minpoly \(z^{4}-3z^{3}+7z^{2}-4z+4 \to \) degree-4 field with quadratic subfield \(\mathbb{Q}(\sqrt{5})\)). Each lock re-run at draft. The \(v1\to v4\) corrections are recorded in \(SCRUTINY_{P}1P3_{r}ound2.md\): v1 called the determinant register \(\mathbb{Z}/2\) (undercount) and the "residue" one bit; \(v2/v3\) over-corrected to \(\mathbb{Z}/6\) (overcount) and mis-stated the order-5 field as \(\mathbb{Q}(\sqrt{41})\) (a discriminant factor); v4 states the register as the three-valued coset \(-\mu _{3}\), the three orientation invariants as distinct, and the order-5 field as degree-4 over \(\mathbb{Q}(\sqrt{5})\). The held-breath quantifier carries the non-degeneracy hypothesis that excludes d \(= 4\)).

\begin{thebibliography}{99}
\bibitem{markov} A.~A. Markoff, \emph{Sur les formes quadratiques binaires ind\'efinies}, Math. Ann. \textbf{15} (1879), 381--406.
\bibitem{goldman} W.~M. Goldman, \emph{The modular group action on real SL(2)-characters of a one-holed torus}, Geom. Topol. \textbf{7} (2003), 443--486.
\bibitem{cl} S.~Cantat and F.~Loray, \emph{Dynamics on character varieties and Malgrange irreducibility of Painlev\'e VI}, Ann. Inst. Fourier \textbf{59} (2009), 2927--2978.
\bibitem{dirichlet} P.~G.~L. Dirichlet, \emph{Vorlesungen \"uber Zahlentheorie}, 1863.
\bibitem{hb} J.~H. Hannay and M.~V. Berry, \emph{Quantization of linear maps on a torus}, Physica D \textbf{1} (1980), 267--290.
\bibitem{cantat} S.~Cantat, \emph{Bers and H\'enon, Painlev\'e and Schr\"odinger}, Duke Math. J. \textbf{149} (2009), 411--460.
\bibitem{bgmw} I.~Biswas, S.~Gupta, M.~Mj, and J.~Whang, \emph{Surface group representations in $SL_2(\mathbb{C})$ with finite mapping class orbits}, Geom. Topol. \textbf{26} (2022), 679--719.
\end{thebibliography}

\end{document}

