% Act II overview: what theta-star is and why it matters

\subsection{Conceptual overview}

In Act~II we treat the phase parameter $\theta$ not as a fixed input of the
Origin Axiom, but as a scan direction across a family of concrete numerical
models. The aim is to test whether the ``non--cancelling remainder''
postulate leaves a quantitative fingerprint at a distinguished value
$\theta_\star$.

Operationally, Act~II is built on three ingredients:
%
\begin{enumerate}
  \item a four--dimensional lattice sum $\Delta\alpha(\theta)$ whose
  $\theta$--dependence we control at high precision;
  \item an effective mass shift $\Delta m^2(\theta)$ in a 1D scalar
  vacuum simulation, extracted from the lattice data and tested against the
  dispersion relation $\omega^2 = k^2 + m_{\rm eff}^2$;
  \item a family of one--dimensional cancellation chains whose residuals
  $|S_N(\theta)|$ expose the presence or absence of a non--cancelling
  remainder when $N$ degrees of freedom are forced to self--cancel.
\end{enumerate}

Throughout Act~II we work with a dimensionless phase parameter
$\theta \in [0,2\pi]$. The numerical evidence below singles out a narrow
interval around
\begin{equation}
  \theta_\star \simeq 2.598 \, ,
\end{equation}
within which three a priori different constructions line up:
%
\begin{itemize}
  \item the 4D lattice observable $\Delta\alpha(\theta)$ develops a
  stable minimum;
  \item the corresponding mass shift $\Delta m^2(\theta_\star)$ produces a
  consistent effective mass in the scalar vacuum dispersion test;
  \item cancellation chains at $\theta_\star$ exhibit anomalously large
  residuals compared to a \emph{theta--agnostic} benchmark
  (the $\phi^\phi$ choice used in Act~I).
\end{itemize}

The rest of this Act is organised as follows.
In \S\ref{subsec:act2-theta-scan} we present the global
$\Delta\alpha(\theta)$ scan, the zoomed minimum and the
high--$R_{\max}$ stability check.
In \S\ref{subsec:act2-scalar-massshift} we translate
$\Delta\alpha(\theta_\star)$ into a mass correction and test it in a
1D scalar vacuum simulation.
In \S\ref{subsec:act2-cancellation-system} we study the
$\theta$--dependence of cancellation chain residuals, focusing on the
$\theta_\star$ anomaly.
Finally, \S\ref{subsec:act2-synthesis} summarises the ``$\theta_\star$
triad'' and interprets it within the Origin Axiom narrative.

