\section{Falsifiability and Stress Tests}
Phase 3 is an exploratory add-on that binds a single fitted parameter $\theta$ (from frozen CKM/PMNS phase targets under a declared ansatz)
to a Phase 2-style injection diagnostic. The core falsifiability questions for this add-on are therefore about (i) target stability,
(ii) ansatz dependence, and (iii) reproducibility invariants.

\subsection{Target-freeze sensitivity}
The CKM and PMNS phase targets are frozen snapshots. A materially different future global fit (or a different ordering assumption for $\delta_{\mathrm{CP}}$)
could shift the best-fit $\theta$ and/or the preferred discrete offset choice. This does not falsify Phase 2,
but it would falsify the Phase 3 \emph{numerical} baseline as stated here if the discrete sweep no longer prefers $b_{\mathrm{PMNS}}=\pi$ under the updated freeze.

\subsection{Discrete hypothesis robustness}
We adopt the Phase 3 baseline hypothesis by minimizing $\chi^2$ over a small discrete set $b_{\mathrm{PMNS}} \in \{0,\pi,-\pi/2,+\pi/2\}$, without fitting $b_{\mathrm{PMNS}}$.
A falsifying outcome for the current baseline would be that modest, pre-declared changes in the discrete candidate set (e.g., adding $\pm \pi/4$)
produce no stable preference or produce a qualitatively different preference without a clear physical rationale.

\subsection{Ansatz dependence}
The mapping $\theta \mapsto (\delta_{\mathrm{CKM}},\delta_{\mathrm{CP}})$ is an explicit modeling choice. If alternative, equally simple one-parameter mappings
(e.g., non-affine monotone maps, or fixed sign flips) fit comparably well or better, then the baseline mapping used here is not uniquely motivated.
This would not falsify the existence of a correlation, but it would weaken any interpretation that relies on a specific functional form.

\subsection{Reproducibility invariants}
All Phase 3 outputs are required to be reproducible by the gate scripts and the paper bundle manifests.
A hard falsification of the Phase 3 computational claim is failure to reproduce the committed bundle artifacts (tables/figures/manifests)
from a clean checkout under the declared environment and commands.
