% main_origin_axiom.tex
% Top-level paper file for the Origin Axiom / theta* programme.

\documentclass[11pt,a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[numbers]{natbib}

% Basic macros
\newcommand{\thetastar}{\theta^\star}
\newcommand{\OmegaLambda}{\Omega_{\Lambda}}
\newcommand{\Omegam}{\Omega_{\mathrm{m}}}
\newcommand{\Hzero}{H_{0}}

\begin{document}

\title{Origin Axiom Act II--III:\\
A $\thetastar$ prior from flavour and a toy effective vacuum for cosmology}

\author{D.\ Mehmetaj and collaborators}
\date{\today}

\maketitle

\begin{abstract}
We explore a toy implementation of the \emph{Origin Axiom}: a programme in
which the vacuum is constrained by a global non-cancelling principle and a
single phase $\thetastar$ simultaneously organises flavour mixing and an
effective vacuum energy. In Act~II we build phenomenological
$\thetastar$--dependent ans\"atze for the leptonic and quark mixing
matrices, fit them to current flavour data, and extract a posterior band
and fiducial value for $\thetastar$. This band is encoded as a simple
configuration file that can be consumed by other codes. In Act~III we
construct lattice toy models of vacuum microstructure, including a
$\thetastar$--dependent microcavity, and we use them to define an
effective-vacuum contribution to a Friedmann--Robertson--Walker (FRW) toy
universe. We compare the resulting age, deceleration parameter, and
Hubble-diagram-style distance relation to standard matter-only and
$\Lambda$CDM benchmarks. The purpose of this paper is methodological: to
demonstrate that a single phase prior derived from flavour can be wired, in
a fully reproducible way, into simple vacuum and cosmology models, and to
make explicit the scaling assumptions required to connect toy microphysics
to observed vacuum energy scales.
\end{abstract}

\tableofcontents

%------------------------------------------------------------
% Act I: Motivation and conceptual backbone
%------------------------------------------------------------
\input{act1_origin_axiom.tex}
% --- Act II: flavour → theta-star prior → vacuum toys
\input{act2_theta_star_section}
\input{act2_frw_vacuum_demo}
\input{act2_frw_from_microcavity}
\input{act2_lambda_scaling_check}

% --- Act III: effective vacuum + FRW observables
\input{act3_effective_vacuum_interface}
\input{act3_frw_observables}
\input{act3_hubble_diagram}
\input{act3_effective_vacuum_band_scan}

% --- Act IV: microstructure upgrades and robustness
\input{act4_microstructure_upgrade}

%------------------------------------------------------------
% Discussion and outlook
%------------------------------------------------------------
\section{Discussion and outlook}
\label{sec:discussion}

We close by summarising the main structural ingredients of the Origin Axiom
programme and outlining concrete next steps. First, we have shown that a
single phase $\thetastar$ can be used as a bridge parameter between
flavour physics and toy models of vacuum microstructure. The $\thetastar$
posterior extracted from flavour fits is encoded in a compact configuration
file and subsequently consumed by independent scripts that implement vacuum
and FRW toy models. This explicit wiring is a small but important step
towards a reproducible ``end-to-end'' story, from flavour data to
cosmological observables, within a controlled toy setting.

Second, we have seen that the non-cancelling principle---implemented here
as a simple clipping rule on coarse-grained amplitudes---naturally
generates persistent vacuum-like contributions whose energy scale can be
matched to the observed cosmological constant by an explicit scaling
assumption. In the present work that scaling is imposed rather than
derived, and the microcavity model is deliberately minimal. Nonetheless,
the resulting effective-vacuum cosmology reproduces the qualitative
features of an accelerating universe, with an age and deceleration
parameter broadly compatible with standard $\Lambda$CDM.

Finally, the limitations of the present toy models suggest a clear roadmap
for future Acts. A more realistic implementation would require:
(i) deriving a non-cancelling constraint from an underlying field theory
or quantum-information-theoretic principle, (ii) constructing microcavity
or defect models tied to specific particle content and interactions,
(iii) integrating the flavour and vacuum sectors into a single dynamical
framework, and (iv) confronting the resulting effective cosmology with
precision data beyond the simple observables considered here. These steps
will be the focus of future work in the Origin Axiom programme.

\nocite{origin_axiom_inprep}
\bibliographystyle{unsrtnat}
\bibliography{origin_axiom}

\end{document}
