% paper/sections/00_abstract.tex
% (Abstract text only; main.tex should wrap this in \begin{abstract}...\end{abstract}.)

We report Phase~II of the \emph{Origin Axiom} program: a reproducible, claim-audited demonstration that a strict global \emph{non-cancellation floor} can be threaded through a minimal chain linking (i) a flavor-inspired phase anchor, (ii) a toy vacuum sector, and (iii) a toy cosmological background, without inducing numerical instability or uncontrolled dynamics.
The axiom is implemented as a hard constraint on a global complex amplitude $A$,
$|A| \ge \varepsilon$ with fixed $\varepsilon>0$, interpreted as forbidding perfect destructive interference (``absolute nothingness'') at the level of the global configuration.

Phase~II is organized around three claims.
(C2.1) \emph{Existence under constraint:} in a controlled lattice field setting, enforcing $|A|\ge\varepsilon$ produces a stable nonzero residual consistent with the intended ``no perfect cancellation'' principle (Fig.~A).
(C2.2) \emph{Robustness and suppression:} the residual and its diagnostics remain well-behaved under systematic variations of $\varepsilon$ and discretization/UV controls, with effects remaining percent-level rather than explosive (Figs.~B--D).
(C2.3) \emph{FRW viability:} mapping the residual into an effective $\Lambda(\theta)$ term yields a modest, smooth perturbation to Friedmann--Robertson--Walker evolution (Fig.~E), demonstrating end-to-end consistency of the toy pipeline.

A single phase parameter $\theta^\star$, treated as an empirical anchor from a separate flavor-motivated fit, is propagated through the chain to test whether a common phase can coherently modulate the residual across modules. We find a small but structured $\theta$-dependence of the vacuum residual (order-percent modulation) and a correspondingly small impact on the FRW trajectory in the explored regime.

We emphasize scope and falsifiability: Phase~II does not derive the axiom from a local action, does not claim a quantitative prediction for the observed cosmological constant, and treats the models as minimal testbeds. All figures are generated from tagged runs with explicit scripts, configurations, and run manifests, enabling full third-party reproduction.