\documentclass[11pt,reqno]{amsart}

% ---------------------------------------------------------------------------
%  The narrow note (post-review rewrite of the flagship's mathematical core).
%  Scope: ONE figure-eight component at SL(4,C) -- the Dehn-filling slope
%  L = -M^4 -- and its completeness, with the metallic trace-map tower as
%  context. Honest-badged; novelty positioned against the projective-geometry
%  and Ptolemy programs; specialist-close pending. The philosophical essay of
%  the flagship is split off and is NOT part of this note.
% ---------------------------------------------------------------------------

\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[margin=1.1in]{geometry}
\usepackage[colorlinks=true,linkcolor=BlueViolet!70!black,citecolor=ForestGreen!60!black,urlcolor=NavyBlue]{hyperref}

\graphicspath{{figures/out/}}

% --- proof-status badges (kept, but used sparingly and never over-stated) ---
\definecolor{stProved}{HTML}{2E7D54}
\definecolor{stNum}{HTML}{1B3A5B}
\definecolor{stStruct}{HTML}{7A4F9E}
\definecolor{stKnown}{HTML}{8C8C8C}
\definecolor{stGated}{HTML}{B6411A}
\newcommand{\badge}[2]{{\setlength{\fboxsep}{2pt}%
  \colorbox{#1}{\textcolor{white}{\sffamily\bfseries\scriptsize #2}}}}
\newcommand{\sProved}{\badge{stProved}{PROVED}}
\newcommand{\sExact}{\badge{stProved}{SYMBOLIC-EXACT}}
\newcommand{\sNum}{\badge{stNum}{NUMERICAL}}
\newcommand{\sKnown}{\badge{stKnown}{KNOWN}}
\newcommand{\sGated}{\badge{stGated}{NEEDS-SPECIALIST}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\SL}{\mathrm{SL}}
\newcommand{\PGL}{\mathrm{PGL}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\chr}{\chi}
\DeclareMathOperator{\Sym}{Sym}

\begin{document}

\title[An SL(4,C) Dehn-filling slope of the figure-eight knot]{An $\SL(4,\C)$ Dehn-filling slope of the figure-eight knot:\\ the relation $L=-M^4$ and its completeness}
\author{Dritëro M.}
\address{Independent researcher, Prishtina.}
\date{\today}

\begin{center}\fbox{\parbox{0.92\linewidth}{\small\textbf{CORRECTION (2026-06-15).} The "component"
framing of this draft is \emph{refuted}: a round-2 deformation-theory computation (verified) shows the
boundary spectrum $\{1,1,\omega,\omega^2\}$ is \emph{not rigid} --- the four-parameter family is a
codimension-one \emph{slice} of a strictly larger component on which $L=-M^4$ fails, not a component. The
honest, stronger result is the rank-stratified degeneration of degree=rank (genuine component at $n=3$,
slice at $n=4$, absent at $n=5$): see \texttt{frontier/B153}. This manuscript is retained as a record;
\textbf{it is superseded and is to be rewritten to the B153 thesis} before any use. Theorem~\ref{thm:relation}
($[A,B]=-\mu^4$ on the family) and the completeness for the fixed spectrum remain correct as
\emph{slice} statements.}}\end{center}

\begin{abstract}
We exhibit an irreducible four-parameter \emph{slice} (\emph{not} a component --- see the correction note
above) of the $\SL(4,\C)$ character variety of the figure-eight knot complement $4_1$ on which the meridian
$\mu$ and longitude $[A,B]$ satisfy the exact matrix identity $[A,B]=-\mu^4$, hence the eigenvalue relation
$L=-M^4$. The slice sits over the
boundary $A$-spectrum $\{1,1,\omega,\omega^2\}$ ($\omega^3=1$); its meridian is of finite order, so the
relation is a \emph{Dehn-filling slope} --- the $\SL(4)$ analogue of Falbel's $\SL(3)$ relations $M^3=L$,
$M^3L=1$, and a distinct object from the boundary-unipotent convex-projective deformations of the
geometric representation studied by Cooper--Long--Thistlethwaite and Ballas. A gauge-stratified primary
decomposition over $\Q(\omega)$ together with an exact finite-field Burnside classification shows this
four-parameter family is the \emph{entire irreducible locus} for that spectrum, so $L=-M^4$ holds there
unconditionally. We frame the result against the metallic trace-map tower (whose adjoint Jacobian factors
over a Dickson catalogue, proved through rank $4$, with a clean $n=2$ base case) and the figure-eight
$A$-polynomial controls at $\SL(2)$ (Cooper--Long) and $\SL(3)$ (Falbel/Heusener--Mu\~noz--Porti). We do
\emph{not} claim a relation ``for all $n$'': the family is established at $n=3$ (known) and $n=4$, and the
$\SL(5)$ status is open (Section~\ref{sec:controls}). An adversarial literature read found no prior
statement of $L=-M^4$ and excludes the convex-projective circle, but the relation is computable by the
extended-Ptolemy $A$-variety algorithm (published only at $n=2$), so a specialist confirmation is the
remaining novelty check. Code and a finite-field certificate accompany the paper.
\end{abstract}

\maketitle

\input{sections/01_intro}
\input{sections/02_tower}
\input{sections/03_dehn_filling}
\input{sections/04_completeness}
\input{sections/05_controls}
\input{sections/06_novelty}
\input{sections/appendix}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
