\section{Reproducibility}
Phase~3 follows the same gate structure as Phases~0--2, with explicit levels and a single
Snakemake entry point. The authoritative, machine-readable reproducibility contract is
documented in \texttt{phase3/REPRODUCIBILITY.md}, but we summarize the structure here.

\paragraph{Levels.}
\begin{itemize}
  \item \textbf{Level A (repo snapshot):} verify the paper and bundle from an existing
        repository snapshot. This level checks that the committed artifacts and the
        bundle manifest are internally consistent. Entry point:
        \texttt{bash scripts/phase3\_gate.sh --level A}.
  \item \textbf{Level B (regenerate artifacts):} rebuild all tables, figures, the paper,
        and the paper bundle from source using Snakemake, then verify the bundle.
        This is the default reproducibility target for external readers:
        \texttt{snakemake -s phase3/workflow/Snakefile -c 1 all}, followed by
        \texttt{bash scripts/phase3\_gate.sh --level B}.
  \item \textbf{Level C (heavy runs):} optional developer mode that can populate
        \texttt{phase3/outputs/runs/} (git-ignored) with additional scans or diagnostics.
        This mode is not required to reproduce any claims in this paper, and is intended
        only for extended exploration:
        \texttt{bash scripts/phase3\_gate.sh --level C}.
\end{itemize}

\paragraph{Bundle and evidence location.}
All artifacts used as evidence for the Phase~3 claims are collected under
\texttt{phase3/outputs/paper\_bundle/}, with a \texttt{run\_index.json} and
\texttt{bundle\_manifest.json} that enumerate inputs and outputs with hashes and sizes.
The canonical PDF for this phase is
\texttt{phase3/artifacts/origin-axiom-phase3.pdf}, built by the Phase~3 Snakemake workflow.
