% phase0/paper/sections/00_abstract.tex

\begin{abstract}
We present Phase~0 of the Origin Axiom program: a foundations and provenance layer that makes subsequent phases auditable, falsifiable, and resistant to accidental numerology.
The program begins from a simple methodological postulate---\emph{global perfect cancellation is forbidden}---and explores its consequences through controlled implementations and domain-to-domain transfers.

Phase~0 contributes four deliverables.
First, we formalize the non-cancellation postulate as a constraint on a specified global complex amplitude $A$ and a strictly positive floor $\varepsilon>0$, and we define \emph{binding} versus \emph{non-binding} regimes to prevent misinterpretation of inert constraints.
Second, we introduce a corridor-based inference protocol for the universal phase parameter $\theta\in[0,2\pi)$: rather than fixing a single $\theta^\star$ early, each phase produces an admissible set (a corridor, possibly disconnected) via explicit pass/fail tests, and the program advances by intersecting these admissible sets across independent layers.
Third, we state phase-by-phase contracts (assumptions, claims permitted, required artifacts, and non-claims) so that each phase can be locked only when its claims are tied to reproducible run manifests and ablation controls.
Fourth, we define a minimal, deterministic ``Phase~0 ledger'' mechanism: future phases emit standardized JSON filter artifacts over $\theta$, which the ledger intersects to produce a versioned corridor history and a human-auditable dashboard after each phase or subphase.

Phase~0 makes no physical claim that the postulate is true in nature, nor does it attempt to derive $\varepsilon$ or a unique $\theta^\star$.
Its purpose is to establish a rigorous scaffold in which later phases can test whether a single phase-like control parameter can coherently survive multiple independent constraints, and whether the corridor narrows toward a discrete or unique value as layers are added.
\end{abstract}