% paper/sections/08_conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

Phase~0 establishes the canonical scaffold for the Origin Axiom program.
We have formalized the non-cancellation postulate as a floor constraint on a phase-specified global complex amplitude $A$ with $\varepsilon>0$, and we have separated binding from non-binding regimes to prevent misinterpretation of inert constraints.
We introduced the corridor method as the program's inference protocol for a universal phase parameter $\theta\in[0,2\pi)$: each phase is required to define explicit pass/fail tests, output an admissible set of $\theta$ values, and propagate progress by intersecting admissible sets across independent layers rather than by fixing $\theta^\star$ prematurely.

We also defined the reproducibility contract that governs the entire series.
Causal claims require ablations; OA-relevance claims require binding certificates; and all headline results require run manifests and full provenance.
Finally, we specified a minimal deterministic ledger mechanism that ingests standardized per-phase JSON filter artifacts to produce a versioned corridor history and a human-auditable dashboard after each phase or subphase.

Phase~0 makes no claim that the Origin Axiom is true in nature, nor does it attempt to derive $\varepsilon$ or a unique $\theta^\star$.
Its purpose is to make the subsequent phases reviewable: if later phases succeed, the program will have earned the right to claim that a single phase-like parameter can survive multiple independent constraints and potentially collapse from a corridor to a discrete or unique value.
If later phases fail, the failure will be legible: it will be attributable to explicit assumptions, invariance limitations, or insufficiently informative tests rather than to hidden drift or untracked choices.

With this scaffold locked, Phase~I and Phase~II can be read as disciplined experiments:
Phase~I validates the implementation and basic binding behavior of OA enforcement, and Phase~II tests whether a constrained residual can be transferred into a stable downstream embedding with controlled dependence on $\theta$.
Subsequent phases may then attempt principled mechanisms, cross-domain predictions, and consolidation only to the extent that the corridor-intersection history supports such claims.