\subsection{Data-driven \texorpdfstring{$\theta_\star$}{theta*} prior from flavor observables}
\label{sec:theta_star_posterior}

In this section we summarise the constraints on the master phase $\theta_\star$ that follow from
charged-current flavor data.  We treat $\theta_\star$ as a single real angle in $[0,2\pi)$
entering the lepton mixing matrix (PMNS) via the Dirac CP phase $\delta_\text{CP}$ and,
in more structured ansätze, also controlling correlated deformations of the three mixing
angles and the two mass-squared splittings.

\paragraph{Ansatz family.}
Within the dedicated phenomenology repository \texttt{origin-axiom-theta-star} we implemented
a small hierarchy of effective ansätze:
\begin{itemize}
  \item a minimal ``sanity'' ansatz \texttt{example\_minimal} in which the PMNS parameters
        are sampled directly around the NuFIT~5.2 central values and no $\theta_\star$ is introduced;
  \item a first non-trivial ansatz \texttt{theta\_star\_delta\_only} in which
        $\delta_\text{CP} \equiv \theta_\star$ while the remaining PMNS parameters are still
        treated as independent;
  \item a structured ansatz \texttt{theta\_star\_v2} in which $\theta_\star$ not only sets
        $\delta_\text{CP}$ but also drives coherent, uncertainty-scaled modulations of the
        three mixing angles and a common fractional rescaling of the mass-squared splittings.
\end{itemize}
For all fits we adopt normal ordering (NO) and NuFIT~5.2 central values and $1\sigma$ errors
as targets for the lepton sector, and standard Wolfenstein parameters for the quark sector.

\paragraph{Sampling and goodness-of-fit.}
We perform flat Monte Carlo scans over the ansatz parameter spaces using three kinds of runs:
(i) PMNS-only runs with \texttt{theta\_star\_delta\_only},
(ii) PMNS-only runs with \texttt{theta\_star\_v2}, and
(iii) joint CKM+PMNS runs with \texttt{theta\_star\_v2}.
Each run generates $\mathcal{O}(10^3$--$10^4)$ samples; for every sample we compute the
total chi-squared $\chi^2_\text{tot} = \chi^2_\text{PMNS} + \chi^2_\text{CKM}$ and
retain only those with $\chi^2_\text{tot} \leq 50$ for the posterior summary.
This cut is deliberately loose: it excludes very poor fits while preserving a wide
region of parameter space around the best solutions.

\paragraph{Global $\theta_\star$ posterior.}
From three representative normal-ordering runs,
\begin{enumerate}
  \item a PMNS-only \texttt{theta\_star\_delta\_only} scan with $N=2000$ samples;
  \item a PMNS-only \texttt{theta\_star\_v2} scan with $N=2000$ samples; and
  \item a joint CKM+PMNS \texttt{theta\_star\_v2} scan with $N=4000$ samples,
\end{enumerate}
we collect all samples that satisfy $\chi^2_\text{tot} \leq 50$ and form a combined
empirical distribution of the input parameter $p_{\theta_\star}$.
The resulting one-dimensional posterior for $\theta_\star$ is summarised by its
16\,\%, 50\,\%, and 84\,\% quantiles,
\begin{equation}
  \theta_\star^\text{(16\%)} \simeq 2.18~\text{rad}, \qquad
  \theta_\star^\text{(50\%)} \simeq 3.63~\text{rad}, \qquad
  \theta_\star^\text{(84\%)} \simeq 5.54~\text{rad} \, .
\end{equation}
In other words, our current flavor-based analysis prefers a broad but non-trivial band
\begin{equation}
  \theta_\star \in [2.18,\,5.54]~\text{rad}
\end{equation}
with a median at $\theta_\star \simeq 3.63~\text{rad}$.  This interval
already represents a significant constraint: not all values in $[0,2\pi)$ are equally
compatible with the observed CKM+PMNS structure, even within the fairly flexible
ansätze explored here.

\paragraph{Working prior for the scalar-universe studies.}
For the purposes of the scalar-universe and cancellation-system analyses that follow
in this work, we adopt this posterior as an \emph{Act~II $\theta_\star$ prior}.
Concretely, we define
\begin{equation}
  \theta_\star^\text{fid} \equiv \theta_\star^\text{(50\%)} \simeq 3.63~\text{rad}
\end{equation}
as a fiducial value for detailed toy-universe simulations, and we treat the
1$\sigma$ band $[2.18,5.54]$~rad as a measure of systematic
uncertainty in the underlying microstructure.
In the subsequent sections, whenever we speak of ``running the cancellation system at
$\theta_\star$'' we mean that the scalar model is evaluated either at
$\theta_\star^\text{fid}$ or at a small number of representative points within this
band.
